\documentclass{article}

\usepackage[english]{babel}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath}
\usepackage{xcolor,graphicx}
\usepackage{hyperref}
\hypersetup{%
  colorlinks=true,
  breaklinks=true,
  linkcolor=blue,
  anchorcolor=black,
  citecolor=brown,
  filecolor=blue,
  menucolor=red,
  urlcolor=blue,
}
\usepackage[authoryear]{natbib}
\usepackage{subcaption}

% -------------
% Some commands
% -------------
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

% -------------

\title{Is Stephen Curry really a guard? New perspective on players typology using functional data analysis}
\author{%
Steven Golovkine\thanks{MACSI, Department of Mathematics and Statistics, University of Limerick, Ireland \href{mailto:steven.golovkine@ul.ie}{steven.golovkine@ul.ie}}
\and
Edward Gunning\thanks{Department of Biostatistics and Epidemiology, University of Pennsylvania, USA \href{mailto:edward.gunning@pennmedicine.upenn.edu}{edward.gunning@pennmedicine.upenn.edu}}
}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We present a novel representation of NBA players' shooting patterns based on Functional Data Analysis (FDA). Each player's charts of made and missed shots are treated as smooth functional data defined over a two-dimensional domain corresponding to the offensive half-court. This continuous representation enables a parsimonious multivariate functional principal components analysis (MFPCA) decomposition, producing a set of common principal component functions that capture the primary modes of variability in shooting patterns, along with player-specific scores that quantify individual deviations from the average behavior. We first interpret the principal component functions to characterize the main sources of variation in shooting tendencies. We then apply k-medoids clustering to the principal component scores to construct a data-driven taxonomy of players. Comparing our empirical clusters to conventional NBA position labels reveals low agreement, suggesting that our shooting-pattern representation might capture aspects of playing style not fully reflected in official designations. The proposed methodology provides a flexible, interpretable, and continuous framework for analyzing player tendencies, with potential applications in coaching, scouting, and historical player or match comparisons.

\end{abstract}

\keywords{Basketball, Clustering, Functional Data Analysis, Sport Analytics}

% MAIN --------
\input{main/introduction}

\input{main/dataset}

\input{main/method}

\input{main/results}

\input{main/discussion}

% APPENDIX ------
\setcounter{equation}{0}
\renewcommand{\theequation}{A.\arabic{equation}}
\begin{appendix}

\input{supp/clustering_results}

\end{appendix}

% ACKNOWLEDGMENT -------
\section*{Acknowledgment}

S. Golovkine is partially supported by Science Foundation Ireland under Grant No. 19/FFP/7002 and co-funded under the European Regional Development Fund.

% REFERENCES ---------
\bibliographystyle{apalike}
\bibliography{./biblio.bib}

\end{document}

