\section*{Associate Editor}

\comment{
The paper demonstrates the use of functional data analysis methods applied to shot chart densities in basketball for investigating common patterns among players. The shot chart is an important object of quantitative analysis in basketball, particularly when investigating spatial features in offensive strategies and performance. While the referees are in agreement that the paper targets a problem of interest, their reports contains several suggestions for revision. Below are some key issues with the paper, some of which overlap with comments from the two referees, that should be carefully and fully addressed in a revision, in addition to other points raised by the referees.
}

\reply{
	LEAVE UNTIL LAST.
}

\comment{
In the introduction, several methods for analyzing shot charts are summarized, including several based on intensity surfaces. The authors cite the requirement of a discretized grid for these methods as the reason for pursuing a density-based analysis, but this is not convincing. All functional data analyses must, in the end, be performed on a discrete grid, and it is not obvious that the estimated (log) intensity surfaces could not also be used as functional data objects. Another key difference between bivariate densities and intensity surfaces is that the latter contain information about shot volume, an aspect completely absent in the density representation. Please provide a more thorough justification for using densities as data objects as opposed to existing approaches.
}

\reply{
	We acknowledge this point, so here we expand on why we chose density surfaces over intensity surfaces:
	\begin{itemize}
		\item \textbf{Intensity surfaces}: mean number of shots expected per unit area. These are \emph{not} normalized. Players who play more games, or just shoot more in general can have different intensities despite having same underlying patterns. \textcolor{red}{I think there are some normalisation, see \cite{millerFactorizedPointProcess2014}, point $3$ of the procedure, but this is not clear how it is done.}
		\item \textbf{Density surfaces}: These are normalized (densities integrate to $1$), hence more suitable for comparing shot profiles, especially across longer time periods. The only drawback is that they can lack information shot volumes.
	\end{itemize}
	Therefore, purely from a behaviour characterization standpoint, we prefer densities because they are normalized and hence allow to make fair comparisons of shooting  \emph{patterns} that are not affected by, e.g., number of games or minutes played. However, we acknowledge that an overall shot volume metric, normalized where appropriate, e.g., total shots per game, could lead to an improved characterization in future work.
	We have added the following sentence to the introduction to reflect this:
	\begin{quote}
		\emph{{\color{blue}``Densities are computed separately for each shot type (made and missed) and each player and are therefore normalized for each individuals total number of each shots.
					We prefer this over an un-normalized representation that contains information about overall shot volume, so that we can make make fair comparisons of shooting
					patterns that are not affected by, e.g., number of games or minutes played."}}
	\end{quote}
	We have also added the following sentences to the discussion:
	\begin{quote}
		\emph{{``\color{blue}Secondly, this work focused on characterising the spatial patterns of made or missed shots. The normalization in the density estimation step, performed separately for the made and missed shots, inherently removes information about total shot volume or shooting efficiency (e.g., percentage chance of success). Future work could include an overall measure of shot volume or shooting efficiency to improve the characterization of players' behaviours."}}
	\end{quote}
	Additionally, the point on FDA methods often requiring discretizations (on a fine grid) is sometimes  correct. However, when a basis representation is used, the calculation is performed using the basis coefficients rather than the discretized objects (see, e.g., \cite{ramsayFunctionalDataAnalysis2005}).
	We argue, however, that there is not a clear consensus from the cited works on how to discretize a (half) basketball court:
	\begin{enumerate}
		\item \cite{millerFactorizedPointProcess2014} and \cite{franksCharacterizingSpatialStructure2015} use one-square-foot tiles for ``computational tractability of LGCP [(log Gaussian Cox process)] inference'' and they ``expect this tile size to capture all interesting spatial variation''.
		\item \cite{yinBayesianNonparametricLearning2022} also use one-square foot tiles (``This rectangle waas divided into $50 \times 35 = 1750$ equally-sized grid boxes of $1ft \times 1ft$'').
		\item \cite{reichSpatialAnalysisBasketball2006} divided the court into an $11 \times 11$ grid based on distance and angle.
		\item \cite{huZeroinflatedPoissonModel2023} partionned the court into $1ft \times 2ft$ blocks.
		\item \cite{fuHoopInSightAnalyzingComparing2024} divided the court ``into a series of contiguous smaller regions for aggregation and convenience''.
		\item \cite{huBayesianGroupLearning2021} partitioned the court to disjoint triangles and used this mesh of discrete sampling locations to estimate a continuous surface in space via interpolation.
		\item \cite{jiaoBayesianMarkedSpatial2021} partitioned the court into $50 x 35$ grid boxes of $1$ft by $1$ft.
		\item \cite{wong-toiJointAnalysisField2022} divided the front court into twelve zones.
		\item \cite{yinAnalysisProfessionalBasketball2023} discretized the court to generate a $25 \times 18$ matrix as the main variable of interest.
	\end{enumerate}
	While discretization is necessary for these approaches and the justification for each type can be understood individually, we argue that the continuous FDA approach removes the subjectivity that may arise.
}

\comment{
While it is true that early work in functional data analysis did include applications of FPCA (or other standard FDA methods) directly to densities \parencite{kneipInferenceDensityFamilies2001,neriniClassifyingDensitiesUsing2007, ramsayFunctionalDataAnalysis2005}, these methods have been superseded in the last decade by non-linear dimension reduction methods due to the fact that density properties are not retained by linear decompositions \parencite{bigotGeodesicPCAWasserstein2017, delicadoDimensionalityReductionWhen2011, petersenFunctionalDataAnalysis2016}. Many if not all of these newer methods can only be applied to univariate densities; however, the authors should include a discussion about the potential pitfalls of FPCA when applied to functional data that, by nature, satisfy nonlinear constraints (positivity and integration to 1). If possible, some attempt to mitigate these issues for the bivariate shot chart densities should be made.
}

\reply{We acknowledge this point -- we have applied FPCA, a conventional linear dimension reduction technique for standard functional data, i.e., curves assumed to belong to the space of $\mathcal{L}_2(\mathcal{T})$ of square-integrable functions on $\mathcal{T}$. However, densities belong to a constrained space -- they are non-negative and integrate to $1$.
	Therefore, reconstructions given by standard FPCA are not guaranteed to satisfy density constraints.
	As such, alternative non-linear approaches have been proposed for modeling univariate densities, e.g., transformations that map the densities to an unconstrained space before doing FPCA \parencite{petersenFunctionalDataAnalysis2016}.
	However, we have chosen to apply standard FPCA, acknowledging its limitations, because:
	\begin{enumerate}
		\item Conventional FPCA is tractable, it is straightforward to describe to a general data-science audience. We can interpret the individual FPCs as a shift of the mean density \parencite{kneipInferenceDensityFamilies2001}. Furthermore, standard computational FPCA methods can be employed to deal with the high-dimensionality of the two-dimensional surfaces.
		\item Although specialized geometrically-aware density approaches can be can give better performance, standard FPCA has been shown to provide a reasonable approximation in many cases \parencite{petersenFunctionalDataAnalysis2016}.
		\item If FPCA approximations of individual densities take negative values, they can be projected back into the density space, by retaining only the positive part and re-normalizing \parencite{petersenFunctionalDataAnalysis2016, daiPrincipalComponentAnalysis2018}.
	\end{enumerate}
	Therefore, we decided to use the standard FPCA approximation rather than develop a new method by extending existing univariate density approaches to the bivariate case.
	We do, however, now acknowledge this as a limitation in our discussion:
	\begin{quote}
		\emph{\color{blue}``We employed standard FPCA which assumes that observations are realisations from an unconstrained space of functions.
			However, densities are constrained because they must be non-negative and integrate to $1$. Reconstructions from standard FPCA do not necessarily satisfy the non-negativity constraint so specialized methods have been developed \parencite{petersenFunctionalDataAnalysis2016}. However, these are for the univariate case. Rather than develop or extend these methodological approaches to the bivariate case, we apply the more tractable, standard FPCA approach while acknowledging its limitations. If required, reconstructions that lie outside the space of densities can be projected back into the density space.''}
	\end{quote}
}

\comment{
The choice of k-medoid clustering is also not discussed in detail. As suggested by the reviewers, please investigate the robustness or variability of your findings to more clustering methods, or provide a justification for this particular approach.
}

\reply{
	We will address all comments regarding the clustering method here. According to \cite{jacquesFunctionalDataClustering2014}, there are mainly four possibilities to cluster functional data: raw-data clustering, $2$-step methods, non-parametric clustering and model-based clustering.
	\begin{itemize}
		\item Raw-data clustering consists in directly clustering the discretization of the surfaces at some points. Because we discretized the density on a grid of $201 \times 201$ points (that would be flatten to vectors of $40401$ points), this is not ideal to cluster the data.
		\item $2$-step methods consist of, first, reducing the dimension of the data, usually by expanding the data into a basis, and second, to perform the clustering on the basis coefficients.
		\item non-parametric clustering consists in using specific distances between surfaces combined with classic clustering algorithms.
		\item model-based clustering assumes probability distributions on the underlying data. As the notion of probability density function is not well defined for functional data, the probability distributions are usually assumed on the coefficients that describe the data.
	\end{itemize}

	According to \cite{zhangReviewClusteringMethods2023}, clustering methods are split into $2$ categories: vector space and function space. (Not sure that we should use this paper.)
}

\comment{
A key drawback of the analysis is its lack of assessment of sampling variability. This is critical both for the reliability of the interpretation of the functional principal components (particularly those with smaller eigenvalues) as well as the clustering results. Without some inferential instrument (e.g., confidence bands), it is impossible to judge the importance of the findings.
}

\reply{
	{\color{purple}\textbf{Steven}: I have added a small comment here, but to satisfy reviewer I think it might be easy to do something simple like compute $5$ bootstrap samples, and show the first 3 PCs from each run and how they don't change much compared to the original.}

	We acknowledge this as a limitation, though it is slightly less relevant than in typical scientific analyses.
	We computed these components based on the whole population of NBA players over a specific period, rather than a sample.
	Therefore, we are not necessarily trying to make inferences about the modes of variability in a larger population.
	That being said, we acknowledge that a population does not have to exist physically -- e.g., the hypothetical notion of a new NBA player not observed or from another time period.
	Therefore, as a simple sensitivity check, we have performed the following {\color{red}\dots \textbf{ADD bootstrap if we do it.}}
}

\comment{
Although the analytical steps are clearly laid out, it is not obvious what the clustering should reflect relative to player positions. Based on the title and some of the conclusions stated by the authors, the underlying assumption seems to be that there is something unexpected or interesting about the lack of agreement between the clustering and designated player positions. However, there seems no obvious reason why these should coincide given different coaching styles and combination of player strengths. Please provide more concrete guidance about how the proposed analytical pipeline can be useful.
}

\reply{
	To address this, we have added the following description and caution to the methods section:
	\begin{quote}
		\emph{{``\color{blue}Moreover, we will compare the estimated clusters to the $k=5$ nominal positions provided by the NBA. We do not expect perfect agreement -- defence, play-making and movement all play a role in addition to shooting. However, because shooting behaviour is a key component of a player's overall game, the positions comparison can help inform the labels we provide to our clusters."}}
	\end{quote}
}

\comment{
Please carefully check the interpretation of each of the functional principal components. For instance, I do not believe it is the case that the first component reflects players who shoot more or less than average. This is because the functions being analyzed contain no information about overall shot volume, but only the distribution of shots. I also do not believe it is possible that the entire first component is positive, since it is known – at least in the univariate case \parencite{kneipInferenceDensityFamilies2001} – that principal components for densities must integrate to zero.
}

\reply{
	Thank you for this comment.
	Indeed, this interpretation was initially phrased incorrectly. We have now replaced this explanation with a correct interpretation, in terms of relative numbers of shots missed and shots made, rather than overall shooting volume.
	Regarding the negativity-positivity comment, we failed to interpret/ describe the small blue (negative) dot directly underneath the basket for the missed shots. An interpretation has been suitably added.
}

\comment{
In addition to potentially comparing with additional clustering methods, further comparisons should be made between the proposed density-based approach and existing intensity-based approaches (functional or non-functional) in order to highlight advantages, if any, of the proposed pipeline.
}

\reply{
	See Comment 1.4 for a discussion on the different clustering approaches.
}
