\section{}

\comment{The current research is characterizing NBA player shooting behavior using functional data analysis. The authors treat made and missed shot locations as bivariate functional data on the offensive half-court and perform a multivariate functional principal component analysis to extract dominant shooting patterns. Then, using k-medoids clustering on the resulting player-specific scores, they generate a data-driven typology of player shooting styles, questioning the validity of traditional NBA positional labels. Several conceptual and technical issues limit both the novelty and the inferential credibility of the results. I therefore recommend a revision before the paper can be considered for publication.}

\reply{
	Thank you for the detailed review of the manuscript. We have addressed each of the concerns point by point below.
}

\comment{While the paper is intended for a quantitative audience, some statistical steps, especially the MFPCA, are described at a relatively informal level. There is no discussion of convergence or uncertainty in the functional principal components.}

\reply{
	Given the applied nature of the paper, we have not presented convergence results for MFPCA. However, a full theoretical treatment of the method can be found in \cite{happMultivariateFunctionalPrincipal2018}.
	While we are primarily concerned with describing the variation in an entire population of NBA players, rather than making inferences using a sample, uncertainty quantification for the MFPCs is not of primary concern either. However, to assess stability/ sensitivity of our approach, we added a simple bootstrap step (see comment $0.6$).
}

\comment{The authors do not evaluate the stability of their clustering. Furthermore, they do not provide external validation (e.g., prediction, test–train splits, comparison with existing clustering approaches). Please consider using metrics such as Adjusted Rand Index or Silhouette scores to assess cluster quality. Alternatively, compare results with simpler clustering methods (e.g., k-means on shooting frequency by location bins) to demonstrate added value.}

\reply{
	Adjusted Rand Index (ARI) and Silhouette coefficients have been computed. ARI between the NBA clusters and the proposed clustrers are close to $0$ but this is expected as we are looking for another typology than usual NBA players. The Silhouete coefficient for the $k$-medoids with equal weights is equal to $0.04$, while the Silhouette coefficient for the $k$-medoids with variance weights is equal to $0.4$. We also computed the Silhouette coefficient of the NBA clustering using the distance matrix used for the $k$-medoids with equal weights ($-0.05$) and using the distance matrix used for the $k$-medoids with variance weights ($-0.1$).
}

\comment{While contrasting the clusters with NBA-defined positions is informative, the resulting disagreement is not deeply analyzed. Are there specific cases where agreement is worse?}

\reply{
	We do highlight that there is agreement for the center category:
	\begin{quote}
		\emph{``The only exception is the NBA-defined `center' category, which aligns with the cluster $1$ in the $k$-medoids results when weighted by the variance explained."}
	\end{quote}
	Since the majority of both clusters and positions (each axis of the confusion matrix) disagree, it is difficult to pinpoint the most disagreement.
}

\comment{The paper uses a 2D KDE with Gaussian kernel and Silverman’s rule but does not specify whether the bandwidth is optimized globally, per player, or per component.}

\reply{
	A different bandwidth is considered for each component (missed and made shots chart) of each player. We have added a sentence to the main text to clarify this.
}

\comment{The model jointly treats made and missed shots, but no attention is given to efficiency metrics (i.e., shooting percentage), which are arguably central to evaluating shot quality.}

\reply{
	We have now added the following to the discussion to acknowledge this point:
	\begin{quote}
		\emph{\color{blue}``Secondly, this work focused on characterising the spatial patterns of made or missed shots. The normalization in the density estimation step, performed separately for the made and missed shots, inherently removes information about total shot volume or shooting efficiency (e.g., percentage chance of success). Future work could include an overall measure of shot volume or shooting efficiency to improve the characterization of players' behaviours."}
	\end{quote}
}


